Leadership across cultures and contexts: an exploratory analysis

### 2020-05-01 leadership_across_cultures_contexts.Rmd
Fixed variables in fucntions cluster discussion section.

Leader qualities formed two top-level clusters, which we identify as a *Prosocial qualities* cluster, and an *Individual qualities* cluster, each of which contained two interpretable and generally supported subclusters. See Figure \@ref(fig:cluster)A.

The *Prosocial qualities* cluster comprised subclusters we labeled, *Cultural conformity* and *Prosocial competencies*, the later of which included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005]. The larger *Prosocial qualities* cluster also correspondes to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership. A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical fromidable*, and *Drug consumption* (although drug use is often an important social activity).

The three most frequent leader quality variables were found in the *Individual qualities* cluster, which had two smaller and strongly supported subclusters: *Social, reproductive, and material success*, which included traits indicative of the benefits of successful leadership, and a more general *Competencies* cluster which included qualities that are not necesssarily prosocial. Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013]. Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme. Several of these measures of success often co-occur. Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:

@@

The *Competencies* subcluster contained a grouping of variables associated with dominance-based leadership, including *Coercive authority* and *Aggressiveness* [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a grouping that included *Knowledgeable/intelligent* and *Experienced/accomplished*, variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return. Interestingly, *Feared* and *Supernatural* clustered more closely with these prestige traits and not among the dominance traits. There is likely a strong connection between shamanism and ritual-based influence and both prestige and dominance styles of leadership. Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership. Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].

@@

The larger organization subcluster included variables such as *Organize cooperation*, *Ritual functions*, and *Military command*. This subcluster aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019]. In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 [p. 127] notes:

>>>>

Leader qualities formed two top-level clusters, which we identify as a "Prosocial qualities" cluster, and an "Individual qualities" cluster, each of which contained two interpretable and generally supported subclusters. See Figure \@ref(fig:cluster)A.

The Prosocial qualities cluster comprised subclusters we labeled, "Cultural conformity" and "Prosocial competencies", the later of which included a mix of positive personal traits, such as generosity and proper behavior, as well as interpersonal skills, having a favorable personality, oratory skill, and charisma, which generally correspond to a charismatic style of leadership that has been identified in industrialized settings and widely discussed in sociology and more recently emphasized in evolutionary psychology models of leadership [@den_hartog_culture_1999; @weber_distribution_1978; @grabo_charismatic_2016; @howell_role_2005]. The larger Prosocial qualities cluster also correspondes to @henrich_big_2015's Big-man mechanism, which suggests positive assortment of followers to prosocial leaders can facilitate the evolution of cooperative social norms and prestige-based leadership. A few variables did not seem to fit prosocial theme, however, such as *Physical health*, *Physical fromidable*, and *Drug consumption* (although drug use is often an important social activity).

The three most frequent leader quality variables were found in the Individual qualities cluster, which had two smaller and strongly supported subclusters: "Social, reproductive, and material success", which included traits indicative of the benefits of successful leadership, and a more general "Competencies" cluster which included qualities that are not necesssarily prosocial. Leaders, probably as a consequence of their ability to provide social benefits and impose social costs, are also characterized by their material, social, and reproductive success (bottom cluster Figure \@ref(fig:cluster)A), which is consistent with much of the theoretical literature focused on associated outcomes of leadership and social hierarchy [e.g., @von_rueden_mens_2016; @bowles_emergence_2010; @glowacki_role_2013]. Again, some variables in this subcluster, such as *Ingroup member* and *Loyalty*, did not quite fit the theme. Several of these measures of success often co-occur. Among the Trobriands, horticulturalists in Papua New Guinea, @ol06-037 [p. 47] describes relationships between polygynous marriages, increased wealth, and widespread social contacts:

@@

The Competencies subcluster contained a grouping of variables associated with dominance-based leadership, including *Coercive authority* and *Aggressiveness* [@tiger_imperial_1971; @barkow_prestige_1975; @barkow_darwin_1989; @henrich_evolution_2001; @cheng_two_2013] as well as a grouping that included *Knowledgeable/intelligent* and *Experienced/accomplished*, variables that have been emphasized by the *Prestige* theory [@henrich_evolution_2001; @cheng_two_2013] and the more general *Computational services* theory [@hagen_leadership_2019; @garfield_evolutionary_2019], and which are among the most strongly supported in the ethnographic record (Figure \@ref(fig:varsupport)), a point to which we shall return. Interestingly, *Feared* and *Supernatural* clustered more closely with the prestige-style traits than with the dominance-style traits. There is likely a strong connection between shamanism and ritual-based influence and both prestige and dominance styles of leadership. Ritual and supernatural capacities may represent a dimension of leader qualities in which dominance and prestige are coupled forming an alternative style of leadership. Artistic performance is also implicated in this style of leadership and ceremonial, artistic performances by shamans are common displays which often incorporate the superhuman abilities of shamans [@singh_cultural_2017].

@@

The larger organization subcluster included variables such as *Organize cooperation*, *Control economics*, and *Military command*. This subcluster aligns with the many evolutionary theories of leadership that emphasize the role of leaders in organizing collective actions [e.g., @hooper_theory_2010; @van_vugt_cognitive_2007; @glowacki_formation_2016; for review, see @garfield_evolutionary_2019]. In the ethnographic record such functions includes controlling and facilitating economic productivity and ensuring collective actions do not overly conflict with individual or household-level goals, such as among the Tukano of the Brazilian Amazon, where @sq19-018 [p. 127] notes:
